---
title: "DataVisualAnalytics_HW1"
author: "Weifeng She"
date: "August 25, 2016"
output: pdf_document
email: wshe1@gatech.edu
---

2. Implement a function that computes the log of the factorial value of an
integer using a for loop. Note that implementing it using log(A)+log(B)+
··· avoids overflow while implementing it as log(A · B · ···) creates an
overflow early on.
```{r}
library(ggplot2)
library(reshape)
logfact1 <- function(n){
   res = 0
   for(i in 1:n){
           res <- res + log(i)
   }
   return(res)
}
```
3. Implement a function that computes the log of the factorial value of an
integer using recursion.

```{r}
logfact2 <- function(n){
   if(n == 1) {
           return(log(1))
   }else{
           return(logfact2(n-1) + log(n))
   }
}

```
4. Using your two implementations of log-factorial in (2) and (3) above, compute
the sum of the log-factorials of the integers 1, 2,...,N for various N
values.
```{r}
# Define the function to calculate the sum using log-factorial
sumLogFact <- function(N, func){
        res = 0
        for(i in 1:N){
                res = res + func(i)
        }
        res
}
NList1 <- seq(from = 0, to = 5000, by = 200)
NList1[[1]] <- 1
# calculate the sum using log-factorial in (2) 
logfact1Value = c()
for(N in NList1){
        result <- sumLogFact(N, logfact1)
        logfact1Value <- c(logfact1Value, result)
        print(result)
}

# calculate the sum using log-factorial in (3) 
options(expressions=500000)
NList2 = seq(from = 0, to = 2200, by = 200)
NList2[[1]] <- 1
logfact2Value = c()
for(N in NList2){
        result <- sumLogFact(N, logfact2)
        logfact2Value <- c(logfact2Value, result)
        print(result)
}

```
5. Compare the execution times of your two implementations for (4) with an
implementation based on the offcial R function lfactorial(n). You
may use the function system.time() to measure execution time. What
are the growth rates of the three implementations as N increases? Use the
command options(expressions=500000) to increase the number of
nested recursions allowed. Compare the timing of the recursion implementation
as much as possible, and continue beyond that for the other two
implementations.
```{r}
logfact1Time <- c()
logfact2Time <- c()
logfact3Time <- c()
for(N in NList1){
        time1 <- system.time(sumLogFact(N, logfact1))[3]
        time3 <- system.time(sumLogFact(N, lfactorial))[3]
        logfact1Time <- c(logfact1Time, time1)
        logfact3Time <- c(logfact3Time, time3)
}

for(N in NList2){
        time2 <- system.time(sumLogFact(N, logfact2))[3]
        logfact2Time <- c(logfact2Time, time2)
}
sup <- replicate(n=length(logfact1Time) -length(logfact2Time), NaN )
logfact2Time <- c(logfact2Time, sup)
timeVector <- cbind(N = NList1, forloop = logfact1Time, recursion = logfact2Time, lfactorial = logfact3Time)
row.names(timeVector) <- NULL
timeVector <- as.data.frame(timeVector)
head(timeVector)
mtimeVector <- melt(timeVector, id = c("N"))

ggplot(mtimeVector, aes(x=N, y=value, color = variable)) + 
  geom_line() +
  geom_point() + 
  ggtitle("Comparing execution time of different functions") +
  ylab("Elapsed time in seconds")

```


From the graph we can see the growth rate of execution time for recursion implementation is expotiential, almost linear for lfactorial, and implementation for for loop is between other two. 


