---
title: "project2.R"
author: "Weifeng She"
date: "11/21/2016"
output: pdf_document
---

1. Use linear regression to predict profit based on all available numeric variables. Graph the train and test MSE as a function of the train set size (averaged over 10 random data partitions as described above)?

```{r}
library(GGally)
library(reshape)
library(ggplot2)
library(tm)
library(caret)
#library(qdap)
library(dplyr)
library(topicmodels)
library(tidyr)
library(reshape2)
library(Matrix)
```

In this dataset numeric variables include: Runtime, Metascore, imdbRating
imdbVotes , tomatoMeter, tomatoRating, tomatoReviews,tomatoFresh, tomatoRotten, tomatoUserMeter, tomatoUserRating, tomatoUserReviews.
Since tomatoMeter = tomatoFresh/(tomatoFresh + tomatoRotten) and tomatoReviews = tomatoFresh + tomatoRotten, both are not used for model building. 

The lm model based on all the numeric variables could get mse at 1.02e+16 for test data.  
```{r}
load("movies_merged")
movies_merged = subset(movies_merged, Type == "movie")
dim(movies_merged) # after subset, there are only 40000 movies left
# remove all the NA rows for Gross and Budget
movies_merged = movies_merged[complete.cases(movies_merged$Gross),]
movies_merged = movies_merged[complete.cases(movies_merged$Budget),]
# create Profit column
movies_merged$Profit <- movies_merged$Gross - movies_merged$Budget
# drop Gross and Budget column
movies_merged$Gross <- NULL
movies_merged$Budget <- NULL # 4558 X 38
movies_merged = movies_merged[movies_merged$Year >= 2000, ] # 3332 X 38
dim(movies_merged) # only 3332 rows left
# print out the column names 
colnames(movies_merged)
# convert factor variable to numeric
movies_merged$Metascore <- as.numeric(as.character(movies_merged$Metascore))
# check how many NAs in each column  
sapply(movies_merged, function(x) sum(is.na(x)))  

# we can see there are no missing value in both Budget and Gross columns
# select the numeric columns 
movies_numeric <- movies_merged[, c(5, 15, 16, 17, 22, 24, 25, 27, 28, 29, 38)]
dim(movies_numeric) # 3332 X 11
head(movies_numeric)

# convert Runtime to numeric value
head(sort(unique(movies_numeric$Runtime)))
tail(sort(unique(movies_numeric$Runtime))) 

# Runtime_num function modified from project 1
Runtime_num <- c()

for(i in seq_along(movies_numeric$Runtime)){
        
        x <- strsplit(movies_numeric$Runtime[i], ' ')[[1]]
        if (length(x) ==  2){ 
                if(x[2] == "min") #3) "xx min"
                {Runtime_num <- c(Runtime_num, suppressWarnings(as.numeric(x[1])))}
                # 4)"xx h"
                else{Runtime_num <- c(Runtime_num, suppressWarnings(as.numeric(x[1])) * 60)}
        }
        # 2) "XX h xx min"
        else if( length((x) == 4)) {
                Runtime_num <- c(Runtime_num, suppressWarnings(as.numeric(x[1])) * 60  + suppressWarnings(as.numeric(x[3])))
        }
        # 1)"N/A"
        else Runtime_num <- c(Runtime_num, NA)
}

movies_numeric$Runtime <- Runtime_num

head(movies_numeric)

# check missing value for each row
sapply(movies_numeric, function(x) sum(is.na(x)))

# remove all the NA rows
# convert missing value to the median of each column
for( i in 1:10){
movies_numeric[, i][is.na(movies_numeric[,i])] <- 
  median(movies_numeric[,i], na.rm = T)
}

# write funtion to calculate mse for train and test dataset
calculate_MSE <- function(dataset, percent){
  
  # splict data to train and test
  sample_size <- floor(percent * nrow(dataset))
  train_index <- sample(seq_len(nrow(dataset)), size = sample_size)
  train <- dataset[train_index,]
  test <- dataset[-train_index,]
  
  # train model
  lm_model <- lm(Profit~., data = train)
  #summary(lm_model)
  # predict train accuracy
  profit_pred_test <- predict(lm_model, newdate = test)
  train_mse <- mean((train$Profit - predict(lm_model, train)) ^ 2)
  test_mse <- mean((test$Profit - predict(lm_model, test)) ^ 2)

  # combine train_mse and test_mse and return it
  c(train_mse,test_mse)  
} 

# write function to do the run calculate_MSE iter times and calculate the mean
calculate_MSE_mean <- function(dataset, iter, percent){
    each_mse <- c(0, 0)
    for(j in 1:iter) {
      mse <- calculate_MSE(dataset, percent)
      each_mse <- each_mse + mse
    }
    each_mse / iter
}

# calculate the mse with different percent of train and test data range from 0.05 to 0.95
final_mse<- vector(, 3)
for(i in 0:18){
  
  percent <- 0.05 + i * 0.05 
  each_mse <- calculate_MSE_mean(movies_numeric, iter = 100, percent = percent)
  
    each_mse <- c(percent, each_mse)
  
  final_mse <- rbind(final_mse, each_mse)
  
}

# remove the first placeholder row  
final_mse <- final_mse[-1, ]
colnames(final_mse) <- c("train_percent", "train", "test")
#head(final_accuracy)
rownames(final_mse) <- NULL

final_mse <- as.data.frame(final_mse)
print(paste("the best mse for train set with only numeric variable is:", min(final_mse$train), sep = " "))
print(paste("the best mse for test set with only numeric variable is:", min(final_mse$test), sep = " "))

final_mse$train_percent <- factor(as.character(final_mse$train_percent))
head(final_mse)
melt_mse <- melt(final_mse, id = "train_percent")

colnames(melt_mse) <- c("train_percentage", "variable", "mse")

ggplot(melt_mse, 
       aes(x = train_percentage, y=mse, color = variable)) + 
  geom_point() + 
  
  ggtitle("Compare model mse with different percentage of training data with only numeric variables") 
  

```

2. Try to improve the prediction quality in (1) as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g.,is_budget_greater_than_3M). Explain which transformations you used and why you chose them. Graph the train and test MSE as a function of the train set size (averaged over 10 random data partitions as described above)? 

The strategy is to create new varibles by taking log, square, cube of each numeric variables, run lm model with all created variables, then only select the variables which contribute significantly to the final model.    

The lm model based on numeric and transformed numeric variables could get mse at 9.2e+15 for test data. 
```{r}
log_val <- function(x) log10(x)
sqr_val <- function(x) x^2
cub_val <- function(x) x^3
func_list <- c(log_val, sqr_val, cub_val)

func_name <- c("log_", "sqr_", "cub_")
col_name <- colnames(movies_numeric) 

# avoid the Profit column
col_name <- col_name[-length(col_name)]
# calculate the mse for lm model with any tranformation of the numeric varibles
# mse_original <- calculate_MSE_mean(movies_numeric, iter = 10, percent = 0.7)
# only select the test mse
# mse_original_test <- mse_original[2]
keeped_columns <- data.frame(matrix(NA, nrow = dim(movies_numeric)[1], ncol= 0))
keeped_column_names <- c()
for (i in seq_along(func_list)){
  # iterate through each transformation function
    func <- func_list[i]
   # iterate through each numeric column
     for (j in seq_along(col_name) ) {
      
       created_col_name <- paste(func_name[i], col_name[j], sep = "") 
       created_col <- func_list[[i]](movies_numeric[, col_name[j]] + 0.01)
       # print(created_col_name)
       # add this created column to movies_numeric dataframe
       #movies_numeric[,created_col_name]<- func_list[[i]](movies_numeric[, col_name[j]] + 0.01)
        #print(head(movies_numeric))
       #mse <- calculate_MSE_mean(movies_numeric, iter = 10, percent = 0.7)
       #print(mse)    
       #mse_test <- mse[2]
          #if((mse_original_test - mse_test)/mse_original_test > 0.05)
           #{ print(created_col_name)
        keeped_columns <- cbind(keeped_columns, created_col)
        keeped_column_names <- c(keeped_column_names, created_col_name)
          #   }
             # remove this newly created column before next iteration
             #movies_numeric <- movies_numeric[, 1:11]
          }         
}

colnames(keeped_columns) <- keeped_column_names
#head(keeped_columns)

# because we cubed each column and some variable will be extreme large, it is necessary to normalize the data. 

# then we normalize the data

preObj <- preProcess(keeped_columns, method=c("center", "scale"))
keeped_columns<- predict(preObj, keeped_columns)
#keeped_columns_1 <- cbind(keeped_columns, Profit = movies_merged$Profit)
#lm_model_keep <- lm(Profit~., data = keeped_columns_1)
#summary(lm_model_keep)

# then we combine all these created columns with numeric data
movies_combined1 <-  cbind(movies_numeric, keeped_columns)
lm_model_1 <- lm(Profit~., data = movies_combined1)
summary(lm_model_1)
# from the summary of the model we only keep the most significat columns

created_numeric <- keeped_columns[,c("log_tomatoRotten","sqr_tomatoRating", "sqr_tomatoFresh", "sqr_tomatoUserReviews", "cub_tomatoRating", "cub_tomatoFresh","cub_tomatoUserMeter","cub_tomatoUserReviews")]
movies_combined2 <-  cbind(movies_numeric, created_numeric)


final_mse<- vector(, 3)
for(i in 0:18){
  
  percent <- 0.05 + i * 0.05 
  each_mse <- calculate_MSE_mean(movies_combined2, iter = 100, percent = percent)
  
  each_mse <- c(percent, each_mse)
  
  final_mse <- rbind(final_mse, each_mse)
  
}

# remove the first placeholder row  
final_mse <- final_mse[-1, ]
colnames(final_mse) <- c("train_percent", "train", "test")
#head(final_accuracy)
rownames(final_mse) <- NULL

final_mse <- as.data.frame(final_mse)

print(paste("the best mse for train set with only numeric variable is:", min(final_mse$train), sep = " "))
print(paste("the best mse for test set with only numeric variable is:", min(final_mse$test), sep = " "))

final_mse$train_percent <- factor(as.character(final_mse$train_percent))
head(final_mse)
melt_mse <- melt(final_mse, id = "train_percent")

colnames(melt_mse) <- c("train_percentage", "variable", "mse")

ggplot(melt_mse, 
       aes(x = train_percentage, y=mse, color = variable)) + 
  geom_point() + 
  
  ggtitle("Compare model mse with diff percent of training data with transformed numeric variables")
```
3. Write code that featurizes genre (can use code from Part-I), actors, directors, and other categorical variables. Explain how you encoded the variables into features.

for categorical variables genre, actors, rate, director, writer, language, country, awards, and production, Website, only rate and production are single variable at each row and we can directly convert them to dummy varibles.     

for categorical variables genre, actors, director, writer, language, country, they all contains multiple varibles for each row. Therefore we could not use one-hot encoding of dummy varibles to convert them into numerical varible. The strategy is to find the terms for each row by converting to document term matrix, find the most abundance terms and only choose these terms to create dummy varible.  

for the Plot column, basically it is just free text. I use topic modeling to automatically classify sets of documents into themes. The algorithm that I used is Latent Dirichlet Allocation(LDA). The basic assumption behind LDA is that each of the documents in a collection consist of a mixture of collection-wide topics. However, in reality we observe only documents and words, not topics – the latter are part of the hidden (or latent) structure of documents. The aim is to infer the latent topic structure given the words and document.  LDA does this by recreating the documents in the corpus by adjusting the relative importance of topics in documents and words in topics iteratively. 

```{r, echo=FALSE}
#length(table(movies_merged$Rated))
#length(table(movies_merged$Production))

# treat with writer column
# select the writer column
writer <- movies_merged$Writer
# convert it to lower case
writer <- tolower(writer)
# inspect data
# head(writer)
# replace everything inside ()
writer <- gsub(" \\([a-z]*\\)","", writer)
# replace space
writer <- gsub(" ","_", writer)
# split string on ",_"
writer <- gsub(",_", " ", writer)
corp <- Corpus(VectorSource(writer))
dtm <- DocumentTermMatrix(corp)
writer_df <- data.frame(as.matrix(dtm)) #5178 rows
# sort the writers by the number of their movies
sort(colSums(writer_df),decreasing = T)[1:10]
# then top 8 is selected
sorted_writer <- sort(colSums(writer_df), decreasing = TRUE)[1:8]
name_list <- names(sorted_writer)
# only subset the top 10 Genres and complete.cases
sub_writer_df <- writer_df[,names(sorted_writer)]
names(sub_writer_df) <- paste("writer_", name_list, sep="")
dim(sub_writer_df) ## 3332 X 8  
#head(sub_writer_df)


# treat with genre
# select the genre column
genre <- movies_merged$Genre
# convert it to lower case
genre <- tolower(genre)
# inspect data
#head(genre)
# split string on ","
genre <- gsub(",", " ", genre)
corp <- Corpus(VectorSource(genre))
dtm <- DocumentTermMatrix(corp)

genre_df <- data.frame(as.matrix(dtm)) #5178 rows
# sort the writers by the number of their movies
sort(colSums(genre_df),decreasing = T)[1:10]
# then top 7 is selected
sorted_genre <- sort(colSums(genre_df), decreasing = TRUE)[1:7]
name_list <- names(sorted_genre)

sub_genre_df <- genre_df[,names(sorted_genre)]
dim(sub_genre_df) ## 3332 
#head(sub_genre_df)

# analyze actors
# select the actors column
actors <- movies_merged$Actors
# convert it to lower case
actors <- tolower(actors)
# inspect data
#head(actors)
# replace space
actors <- gsub(" ","_", actors)
# split string on ",_"
actors <- gsub(",_", " ", actors)
corp <- Corpus(VectorSource(actors))
dtm <- DocumentTermMatrix(corp)
actors_df <- data.frame(as.matrix(dtm)) #5178 rows
# sort the actors by the number of their movies
sort(colSums(actors_df),decreasing = T)[1:15]
# then top 8 is selected
sorted_actors <- sort(colSums(actors_df), decreasing = TRUE)[1:8]
name_list <- names(sorted_actors)
# only subset the top 8 actors
sub_actors_df <- actors_df[,names(sorted_actors)]
names(sub_actors_df) <- paste("actor_", name_list, sep="")

dim(sub_actors_df) ## 3332X8
#head(sub_actors_df)

# analysize director
# select the Director column
director <- movies_merged$Director
# convert it to lower case
director <- tolower(director)
# inspect data
#head(director, 10)
# replace space
director <- gsub(" ","_", director)
# split string on ",_"
director <- gsub(",_", " ", director)
corp <- Corpus(VectorSource(director))
dtm <- DocumentTermMatrix(corp)
director_df <- data.frame(as.matrix(dtm)) #3332 X 2081 
# sort the director by the number of their movies
sort(colSums(director_df),decreasing = T)[1:15]
# then top 7 is selected
sorted_director <- sort(colSums(director_df), decreasing = TRUE)[1:7]
name_list <- names(sorted_director)
# only subset the top 7 directors and complete.cases
sub_director_df <- director_df[,names(sorted_director)]
names(sub_director_df) <- paste("directo_", name_list, sep="")

dim(sub_director_df) ## 3332 
#head(sub_director_df)

# analysize language
# select the Language column
language <- movies_merged$Language
# convert it to lower case
language <- tolower(language)
# inspect data
#head(language, 10)
# replace comma with space
language <- gsub(","," ", language)
corp <- Corpus(VectorSource(language))
dtm <- DocumentTermMatrix(corp)
language_df <- data.frame(as.matrix(dtm)) #3332 X 133
# sort the language by the number of related movies
sort(colSums(language_df),decreasing = T)[1:15]
# then top 9 is selected
sorted_language <- sort(colSums(language_df), decreasing = TRUE)[1:9]
name_list <- names(sorted_language)
# only subset the top 9 languages are selected
sub_language_df <- language_df[,names(sorted_language)] # 9 column
dim(sub_language_df) ## 3332 
#head(sub_language_df)


# analysize country
# select the country column
country <- movies_merged$Country
# convert it to lower case
country <- tolower(country)
# inspect data
#head(country, 10)
# replace comma with space
country <- gsub(","," ", country)
corp <- Corpus(VectorSource(country))
dtm <- DocumentTermMatrix(corp)
country_df <- data.frame(as.matrix(dtm)) #3332 X 95
# sort the country by the number of related movies
sort(colSums(country_df),decreasing = T)[1:15]
# then top 6 is selected
sorted_country <- sort(colSums(country_df), decreasing = TRUE)[1:6]
name_list <- names(sorted_country)
# only subset the top 6 countries are selected
sub_country_df <- country_df[,names(sorted_country)] # 6 column
dim(sub_country_df) ## 3332 
#head(sub_country_df)


# deal with the awards 

#since rewards column has two types of variables: awards and normination and it has inherent order, we can create the level based on number of reward. First we count total number of awards and normination, then we cut it into 4 ranges according to the quantile of total number of awards and normination and set it to one of the 4 levels (part of the codes were taken from my project1)

awards_normination <- c()

for( i in seq_along(movies_merged$Awards)){
        if(movies_merged$Awards[i] == "N/A"){
                awards_normination <- c(awards_normination, 0)
        }else{
                x <- movies_merged$Awards[i]
                temp <- gregexpr("[0-9]+", x) # find the numbers with any number of digits
                # extract the awards and normination and sum them  
              total_awards_normination  <- sum(as.numeric(unlist(regmatches(x, temp))))
  
         awards_normination <- c(awards_normination, total_awards_normination)
        }
}

# Then we can use quantile function find the distriution of the awards
quantile(awards_normination)
awards_df <- data.frame(matrix(0, nrow = nrow(movies_merged), ncol = 4))
colnames(awards_df) <- c("award_l_1", "award_l_2", "award_l_3", "award_l_4")
for(i in seq_along(awards_normination)){
  if (awards_normination[i] > 16){
    awards_df[i, 4] <-  1
  }else if(awards_normination[i] > 5){
    awards_df[i, 3] <- 1
  }else if(awards_normination[i] > 1){
    awards_df[i, 2] <- 1
  }else{awards_df[i, 1] <- 1}
}
dim(awards_df)

# deal with rate 
table(movies_merged$Rated)
#we can see there are only 11 types of rate varibles. And only 3 types(PG, PG-13, R) are totally dominant, so we can set 3 levels for rate.  
rate_df <- data.frame(matrix(0, nrow = nrow(movies_merged), ncol = 3))
colnames(rate_df) <- c("rate_PG", "rate_PG_13", "rate_R")
for(i in seq_along(movies_merged$Rated)){
  if (movies_merged$Rated[i] == "PG"){
    rate_df$rate_PG[i] = 1
  }else if(movies_merged$Rated[i] == "PG-13"){
    rate_df$rate_PG_13[i] =1
  }else if(movies_merged$Rated[i] == "R"){
    rate_df$rate_R[i] =1
  }
}
# check the convertion is corrected or not
colSums(rate_df)
dim(rate_df)


# deal with Production
length(sort(table(movies_merged$Production), decreasing = T))
# we can see there are 562 productions. Then we can take a look of the very top ones 
sort(table(movies_merged$Production), decreasing = T)[1:15]
# then I decided to choose the production which has at least 100 movies, that will give me 5 virables. 
production_df <- data.frame(matrix(0, nrow = nrow(movies_merged), ncol = 5))
colnames(production_df) <- c("pro_warner", "pro_universal", "pro_20th_cen", "pro_paramount", "pro_sony")
for(i in seq_along(movies_merged$Production)){
  if ( movies_merged$Production[i] == "Warner Bros. Pictures" ){
    production_df$pro_warner[i] <-  1
  }else if(movies_merged$Production[i] == "Universal Pictures"){
    production_df$pro_universal[i] <- 1
  }else if(movies_merged$Production[i] == "20th Century Fox"){
    production_df$pro_20th_cen[i] <- 1
  }else if(movies_merged$Production[i] == "Paramount Pictures"){
    production_df$pro_paramount[i] <- 1
}else if(movies_merged$Production[i] == "Sony Pictures"){
    production_df$pro_sony[i] <- 1
}
}
# check the convertion is corrected or not
colSums(production_df)
dim(production_df)

# treat with Plot

# create a corpus from vector
docs <- Corpus(VectorSource(movies_merged$Plot))
# start processing
# transfer to lower case
docs <- tm_map(docs, content_transformer(tolower))
# remove potentially problematic symbols
toSpace <- content_transformer(function(x, pattern) { return (gsub(pattern, " ", x))})
  docs <- tm_map(docs, toSpace, "-")
  docs <- tm_map(docs, toSpace, "'")
  # remove punctuation               
  docs <- tm_map(docs, removePunctuation)
  # remove digits              
  docs <- tm_map(docs, removeNumbers)
  # remove stopwords
  docs <- tm_map(docs, removeWords,stopwords("english"))
  # remove whitespace
  docs <- tm_map(docs, stripWhitespace)
  # stem document
  docs <- tm_map(docs, stemDocument)
  # check every row
  #as.character(docs[1])
  #Create document-term matrix
  dtm <- DocumentTermMatrix(docs)
#dtm1 <- dtm[, colSums(se)]
  # inspect the term frequency matrix
  dim(as.matrix(dtm))
# find which row.sum equals to 0
raw.sum = apply(dtm, 1, FUN= sum)

zero_rows = which(raw.sum== 0)

length(zero_rows)
dtm1 <- as.matrix(dtm)
#dtm1[1:5, 1:5]  
# set a random column with 1 for all 0 rows to avoid error
for(i in 1:dim(dtm1)[1]){
  if (raw.sum[i] == 0){
    col = sample(dim(dtm1)[2], 1)
    dtm1[i, col] <- 1
  }
}
#collapse matrix by summing over columns
  freq <- colSums(as.matrix(dtm1))
  #length should be total number of terms
  print(paste("The size of the vocabulary is : ",length(freq)))
  #create sort order (descending)
  ord <- order(freq,decreasing=TRUE)
  #List all terms in decreasing order of freq and write to disk
print("The top frequencied words: ")  
freq[ord][1:10]
  
  # start the topicmodeling
  #Set parameters for Gibbs sampling
  burnin <- 4000 # the steps of the walk discarded
  iter <- 2000 # num of iterations
  thin <- 500 # taking every 500th iteration for futher use to avoid correlation btw samples
  seed <-list(2003,5,63,100001,765) # set seed
  nstart <- 5 # use 5 diff starting points for 5 independent runs
  best <- TRUE # instructs the algorithm to return results of the run with the highest posterior probability
  
  # Number of topics
  k <- 5
 #Run LDA using Gibbs sampling
ldaOut <-LDA(dtm1,k, method="Gibbs", control=list(nstart=nstart, seed = seed, best=best, burnin = burnin, iter = iter, thin=thin))


  # write out results
  # docs to topics
  ldaOut.topics <- as.matrix(topics(ldaOut))
  dim(ldaOut.topics)
  
  #top 6 terms in each topic
  ldaOut.terms <- as.matrix(terms(ldaOut,6))
print(ldaOut.terms)
  #write.csv(ldaOut.terms,file=paste("LDAGibbs",k,"TopicsToTerms.csv""))
  #dim(ldaOut)
  #probabilities associated with each topic assignment
  topicProbabilities <- as.data.frame(ldaOut@gamma)
  dim(topicProbabilities)
  colnames(topicProbabilities) <- c("top_1", "top_2", "top_3", "top_4", "top_5")
  #Find relative importance of top 2 topics
  #topic1ToTopic2 <- lapply(1:nrow(dtm),function(x)
  #  sort(topicProbabilities[x,])[k]/sort(topicProbabilities[x,])[k-1])
#}


#ggplot(coef_df[coef_df$coef > 10,], aes(reorder(variables,coef), coef)) +
#geom_bar(stat = "identity", fill = "#9999CC") +
#coord_flip()  +
#xlab("Director")

```

4. Use linear regression to predict profit based on all available non-numeric variables (using the transformations in (3). Graph the train and test MSE as a function of the train set size (averaged over 10 random data partitions as described above)?

First I build a lm model based on all the created categorical variables from question 3. Then select the significant variables. The lm model based on selecyted transformed categorical variables could get mse at 1.3e+16 for test data. 

```{r}
# combine all the variables created in question 3
movies_categorical <- cbind(sub_writer_df, sub_genre_df, sub_actors_df, sub_director_df, sub_language_df, sub_country_df, awards_df, rate_df, production_df, topicProbabilities, Profit = movies_merged$Profit)

dim(movies_categorical) # 3332 X 63

lm_model_2 <- lm(Profit~., data = movies_categorical)
summary(lm_model_2)

# then I selected all the columns which are significant on the lm model. 
created_categorical <- movies_categorical[,c("writer_david_s._goyer","drama", "action", "adventure", "actor_johnny_depp", "usa",  "award_l_1", "award_l_2", "award_l_3", "pro_warner",  "pro_universal", "pro_20th_cen","top_1", "Profit")]

final_mse<- vector(, 3)
for(i in 0:18){
  
  percent <- 0.05 + i * 0.05 
  each_mse <- calculate_MSE_mean(created_categorical, iter = 100, percent = percent)
  
  each_mse <- c(percent, each_mse)
  
  final_mse <- rbind(final_mse, each_mse)
  
}

# remove the first placeholder row  
final_mse <- final_mse[-1, ]
colnames(final_mse) <- c("train_percent", "train", "test")
#head(final_accuracy)
rownames(final_mse) <- NULL

final_mse <- as.data.frame(final_mse)
print(paste("the best mse for train set with transformed categorical variable is:", min(final_mse$train), sep = " "))
print(paste("the best mse for test set with only transformed categorical variable is:", min(final_mse$test), sep = " "))


final_mse$train_percent <- factor(as.character(final_mse$train_percent))
head(final_mse)
melt_mse <- melt(final_mse, id = "train_percent")

colnames(melt_mse) <- c("train_percentage", "variable", "mse")

ggplot(melt_mse, 
       aes(x = train_percentage, y=mse, color = variable)) + 
  geom_point() + 
  
  ggtitle("Compare model mse with different percentage of training data on non-numerical varible")

```

5. Try to improve the prediction quality in (1) as much as possible by using both numeric and non- numeric variables as well as creating additional transformed features including interaction features (for example is_genre_comedy x is_budget_greater_than_3M). Explain which transformations you used and why you chose them. Graph the train and test MSE as a function of the train set size (averaged over 10 random data partitions as described above)?

I first combine 1) all the numeric variables from original data, 2) transfromed numeric varibles from question 2, and 3) transformed categorical variables from question 3 and 4. Then build a lm model to remove the non-significant variables. Based on these variables, the mse of this model for test data can reach 8.9e+15. Then I create variables for all possible interactions btween numerical variables and categorical variables. Then build a lm model to remove the non-significant variables. Based on these variables containing selected interaction variables, the mse for test data can reach 8.7e+15.    

```{r}
movies_combined2$Profit <- NULL
movies_combined3 <- cbind(movies_combined2, created_categorical)
lm_model_3 <- lm(Profit~., data = movies_combined3)
summary(lm_model_3)

# from model summary, we can see some varialbes did not contribute, we could drop it. 
movies_combined4 <- subset(movies_combined3, select = -c(actor_johnny_depp, usa, pro_universal))

lm_model_4 <- lm(Profit~., data = movies_combined4)
summary(lm_model_4)


final_mse<- vector(, 3)
for(i in 0:18){
  
  percent <- 0.05 + i * 0.05 
  each_mse <- calculate_MSE_mean( movies_combined4, iter = 100, percent = percent)
  
  each_mse <- c(percent, each_mse)
  
  final_mse <- rbind(final_mse, each_mse)
  
}

# remove the first placeholder row  
final_mse <- final_mse[-1, ]
colnames(final_mse) <- c("train_percent", "train", "test")
#head(final_accuracy)
rownames(final_mse) <- NULL

final_mse <- as.data.frame(final_mse)
print(paste("the best mse for train set with combined numeric and transformed categorical variable is:", min(final_mse$train), sep = " "))
print(paste("the best mse for test set with combined numeric and  transformed categorical variable is:", min(final_mse$test), sep = " "))


final_mse$train_percent <- factor(as.character(final_mse$train_percent))
#head(final_mse)
melt_mse <- melt(final_mse, id = "train_percent")

colnames(melt_mse) <- c("train_percentage", "variable", "mse")

ggplot(melt_mse, 
       aes(x = train_percentage, y=mse, color = variable)) + 
  geom_point() + 
  
  ggtitle("Compare model mse with different percentage of training data on numerical and categorical variables") 


# then I tried to further improve the mse by creating the interaction between numeric variables and categorical variables. 
numeric_v <- 1:18
categorical_v <- 19:28
inter_data <- data.frame(matrix(0, nrow = dim(movies_numeric)[1], ncol= 0))
inter_name <- c()
for(i in numeric_v){
  name1 <- colnames(movies_combined4)[i]
  
  for(j in categorical_v){
    name2 <- colnames(movies_combined4)[j]
    new_name <- paste("inter", name1, name2, sep = "_")
    inter_name <- c(inter_name, new_name)
    new_col = movies_combined4[i] * movies_combined4[j]
    inter_data <- cbind(inter_data, new_col)
  }
}
colnames(inter_data) <- inter_name

movies_combined5 <- cbind(movies_combined4, inter_data)
# run a lm model on it
lm_model_5 <- lm(Profit~., data = movies_combined5)
# then only select the significant columns
movies_combined6 <- movies_combined5[, (summary(lm_model_5)$coefficients[, 4] < 0.05)]
movies_combined6 <- cbind(movies_combined6, Profit = movies_merged$Profit)


final_mse<- vector(, 3)
for(i in 0:18){
  
  percent <- 0.05 + i * 0.05 
  each_mse <- calculate_MSE_mean( movies_combined6, iter = 100, percent = percent)
  
  each_mse <- c(percent, each_mse)
  
  final_mse <- rbind(final_mse, each_mse)
  
}

# remove the first placeholder row  
final_mse <- final_mse[-1, ]
colnames(final_mse) <- c("train_percent", "train", "test")
#head(final_accuracy)
rownames(final_mse) <- NULL

final_mse <- as.data.frame(final_mse)
print(paste("the best mse for train set with combined numeric and transformed categorical variable is:", min(final_mse$train), sep = " "))
print(paste("the best mse for test set with combined numeric and  transformed categorical variable is:", min(final_mse$test), sep = " "))


final_mse$train_percent <- factor(as.character(final_mse$train_percent))
#head(final_mse)
melt_mse <- melt(final_mse, id = "train_percent")

colnames(melt_mse) <- c("train_percentage", "variable", "mse")

ggplot(melt_mse, 
       aes(x = train_percentage, y=mse, color = variable)) + 
  geom_point() + 
  
  ggtitle("Compare model mse with numerical, categorical and interacted variables") 

```
